---
title: "Steak preferences in th USA"
subtitle: "Summary Report for the  `r params$region` Region"
author: "Author : D.PIERROT" 
date: "`r format(Sys.Date(), '%A, %B, %d, %Y')`" 
output: html_document
params:
  region: 
    label: "Region:"
    value: Mountain
    input: select
    choices: ["East North Central","East South Central", 
              "Middle Atlantic","Mountain", "New England", 
              "Pacific","South Atlantic", "West North Central",
              "West South Central"]

    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
#set any missing Nas int ables to blank
options(knitr.kable.NA='')

library(fivethirtyeight)
library(tidyverse)
library(knitr)
library(kableExtra)
library(ggthemes)

data("steak_survey", package="fivethirtyeight")
 sdat <- na.omit(steak_survey) %>%
  filter(region==params$region)
```

## Background

In May 2014, Walt Hickey at
[538.com](http://fivethirtyeight.com/) published an
article entitled
["How Americans Like Their Steak"](https://fivethirtyeight.com/features/how-americans-like-their-steak/). This article utilized survey data collected from 550 people which asked questions related to various risky activities, such as whether they speed in traffic, if they smoke or go skydiving, and if they prefer a riskier lottery. The survey also asked if they eat steak and if so, how they prefer their steak.

## Purpose

The "steak survey" dataset that was used for the ["How
Americans Like Their
Steak"](https://fivethirtyeight.com/features/how-americans-like-their-steak/) article, is available through
the R package [`fivethirtyeight`](https://cran.r-
project.org/web/packages/fivethirtyeight/index.html).
Using this dataset, this article summarizes the steak
preferences for the _**`r params$region`**_ Region of the
United States by gender.

For our summary presented here, a subset of the original steak survey was extracted to only include responses with no missing data and only included respondents from the _**`r params$region`**_ region of the US. The data subset used for our summary had `r nrow(sdat)` respondents.


## Demographic Summary Tables

The age categories of the `r nrow(sdat)` survey
respondents were:

```{r}
# create table of the age categories
tb <- sdat %>%
  select(age) %>%
  table() %>%
  prop.table()*100

# convert table to data frame
dt <- as.data.frame(tb)

# use kable from knitr package to make table
# and use kable styling from kableExtra package
knitr::kable(dt, format="html",
             col.names=c("Ages","%"),
             digits=2,
             caption="Ages of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options ="striped",
                            full_width=FALSE)
```


## Education Summary Tables


The education levels ot the `r nrow(sdat)' survey respondents were:

```{r}
# create table of the age categories
tb <- sdat %>%
  select(educ) %>%
  table() %>%
  prop.table()*100

# convert table to data frame
dt <- as.data.frame(tb)

# use kable from knitr package to make table
# and use kable styling from kableExtra package
knitr::kable(dt, format="html",
             col.names=c("Education ","%"),
             digits=2,
             caption="Ages of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options ="striped",
                            full_width=FALSE)
```


The income levels ot the `r nrow(sdat)' survey respondents were:



```{r}
# create table of the age categories
tb <- sdat %>%
  select(hhold_income) %>%
  table() %>%
  prop.table()*100

# convert table to data frame
dt <- as.data.frame(tb)

# use kable from knitr package to make table
# and use kable styling from kableExtra package
knitr::kable(dt, format="html",
             col.names=c("Education ","%"),
             digits=2,
             caption="Income of Survey Respondents") %>%
  kableExtra::kable_styling(bootstrap_options ="striped",
                            full_width=FALSE)
```


```{r}
# create table summary for ages
tb1 <-sdat %>% 
  select(age) %>%
  table() %>%
  prop.table()*100

# create table summary for education
tb2 <-sdat %>% 
  select(educ) %>%
  table() %>%
  prop.table()*100

# create table summary for income
tb3 <-sdat %>% 
  select(hhold_income) %>%
  table() %>%
  prop.table()*100

# convert all tables to data frames
tb1df <-as.data.frame(tb1) 
tb2df <-as.data.frame(tb2)
tb3df <-as.data.frame(tb3)

# merge 1st 2 data frames together
mtb <-merge(data.frame(tb1df, row.names = NULL),
            data.frame(tb2df, row.names = NULL),
            by=0, all=TRUE)[-1]

# merge result with 3rd data frame
mtb2 <-merge(data.frame(mtb, row.names = NULL),
             data.frame(tb3df, row.names = NULL),
             by=0, all=TRUE)[-1]

# use the final data frame
# make into a table with kable
# add styling with kableExtra
# add header with labels spanning 2 columns each
mtb2 %>%
 knitr::kable(format="html",
              col.names=c("Category","%","Category","%","Category","%"),
              digits=2,
              caption="Demographics of Survey Respondents") %>%
  kableExtra::kable_styling(c("striped","bordered"),
                            full_width=FALSE) %>%
  add_header_above(c("Ages"=2,"Education"=2,"Income"=2))
```



```{r}
ggplot(sdat, aes(x = steak_prep, fill = female)) +
  geom_bar(position="dodge",colour="black") +
  ggtitle(paste0("Steak Preference by Gender: ",
          params$region," Region")) +
  xlab("Steak Preparation Preference") +
  ylab("Number of Respondents") +
  scale_fill_manual(values=c("skyblue","palevioletred"),
                    name="Gender",
                    breaks=c(FALSE,TRUE),
                    labels=c("Male", "Female")) +
  theme_fivethirtyeight()
```

